### 2021-05-02 Paper/Methodology.tex
paused work for April 28
big push in latex

As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns. In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model". This analysis method is built on the Fama French three factor model and relies on the two-pass sort with an assigned series of beta decile and size decile-based portfolios. Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.

For this analysis, I sort each security into decile groups for beta and size (based on market cap). As firms' size and volatility are not constant functions, each firm is assigned a decile for each month depending on the most recent decile split.  Each security will be assigned a decile for beta and size according to the CRSP /NYSE American beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size. After sorting each of the securities by beta and size decile, 300,264 of the original 830,889 data points had matches for both beta and size portfolios. To estimate the factor sensitivities base on the described decile groups, I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.

@@

Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort. The second pass of the two-pass sort is described in the equation  \eqref{second_pass} and is used to obtain the risk premia. To consolidate the risk premium across time, a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.

\begin{equation}
    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
    \label{second_pass}
\end{equation}

>>>>

As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns. In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model". This analysis method is built on the Fama French three-factor model and relies on the two-pass sort. Like the section \ref{Data} on data, this section is broken up into multiple parts. First, I will explain the methodology for assessing the sensitivity and risk premium for the ESG factor created from the MSCI ESG index minus the petroleum index. Next, I will discuss the methodology for MSCI ESG rating ESG factors. I will conclude with a brief description of the newey west errors and the setup of assessing statistical significance.

For the first part of this study, I sort each security into decile groups based on beta and size. The size and beta deciles ratings are provided by CRSP and are merged on the individual firm level for each month of data. Next, I conducted the first pass for the two-pass sort by running the regression of expected returns on market risk premia the four Fama French factors and the EMP factor as described in equation \eqref{six_factor_basic_reg}. This regression was run twice: once on the beta-ranked portfolios and again on the size-ranked portfolios.

@@

Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort. The second pass of the two-pass sort is described in the equation  \eqref{second_pass}. The second pass describes the risk premia for each of the six factors in the mode. This second pass was performed for both beta and size-based portfolios.

\begin{equation}
    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
    \label{second_pass}
\end{equation}\\

The second set of regressions is focused on establishing a link between MSCI ESG ratings and ESG risk factor sensitivity. Similar to the first set of regressions, I use a two-pass sort methodology using both size and beta ranked portfolios. To create the ESG factor for this set of analyses, I used the all industry ESG factor described in Section \ref{Data}. The process of implementing the regression follows the same basic form as the one for the MSCI ESG index and petroleum index ESG factor by first running a pass on the portfolio by size and following that with a cross-sectional analysis to quantify the risk premium.

Finally, in determining the statistical significance of each of the coefficients, it is important to discuss the newly west error. The newly west error is a replacement for standard error in a traditional OLS regression. the advantage of the newly west is an ability to account for serial correlation within the model

### 2021-05-02 Paper/Results.tex
big push in latex

Table 3 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The column axis is the distinct size deciles where 1 is the smallest market cap companies, and 10 is the decline with the largest market cap companies. The row axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. Finally, the intercept is the excess returns for the given decile. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor. In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts. Table 4 complements table 3 with a breakdown of significance for each coefficient in table 3. The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level. It is important to note that all of the other Fama French factors are also less significant In the largest and small portfolios. This trend might indicate that the data might have pointed at far ends which do not behave like the other deciles.  Table 6 demonstrates the risk premia for each of the factors using the second step in the two-pass sort. All of the Fama French factors in the size decile two-pass sort have positive risk premia. The EMP factor large than all of the other factors, with a coefficient of approximately  16.8\%

Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest beta. Similarly, beta decile 10 is the portfolio of companies with the largest beta values.  Similar to the size decile, the EMP factor exhibits a negative relation with regard to its sensitivity. However, unlike the size decile analysis, there does not appear to be a significant relationship between the beta decile and the factor's sensitivity. Although the most significant impact is on the lowest decile, the least impact is experienced on the 6th decile portfolio, and the largest decile is close to the middle in terms of sensitivity. Similar to the size decile analysis, the EMP coefficients are all very significant, with every decile having significance at the 5\% level. When considering the second phase of the two-pass sort, the similarities to the size decile continue wherein each of the Fama French five factors exhibit positive risk premia, and the EMP also exhibits a positive relationship. The EMP factor again has a high coefficient with approximately 15.8\%.

>>>>

Table 2 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap. The column axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. Finally, the intercept is the excess returns for the given decile. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor. In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is largest for small companies. The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios. However, several of the Fama French factors are also not significant for the largest and smallest portfolios, which may indicate that these tail portfolios might be behaving differently due to their composition rather than a reflection of the factor.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/size_output_inital}\\
    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size. the row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
    \paperspacingwide
\end{center}

Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for Size decile portfolios. The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level. The other two factors have a negative sign and are not statistically significant at the 10\% level.  EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in returns. Although this is by far the smallest coefficient of the four statistically significant returns, this term is still economically meaningful and represents a risk premia for investing in high ESG firms. Overall the model appears to be a fairly good fit with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well fit model \hl{cite this paper}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2}\\
    \textbf{Table 3:} second pass results for the size and beta decile portfolios\\
    \paperspacingwide
\end{center}

Table 4 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the row axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest market beta's. Similarly, beta decile 10 is the portfolio of companies with the largest market beta values.  Unlike the size decile portfolios, the beta decile portfolios do not have consistently negative relations with respect to decile. Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor. \hl{why might this be}. Similar to the size decile analysis, the EMP coefficients are all very significant, with every portfolio having significance at the 5\% level.

When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB CMA and EMP exhibit positive risk premia. The other two factors remain statistically insignificant, and the MHL factor changed signs and now has a positive relationship. The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in returns for a high ESG company. This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude. The model is also a good fit with an $R^2$ value of 86.0%

\begin{center}
    \paperspacingnarrow
    \input{Code/data/beta_output_inital}\\
    \textbf{Table 4:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
    \paperspacingwide
\end{center}

Moving on to the MSCI based portfolio, a two-pass sort was also used to assess the sensitivity and risk premia on the ESG portfolio on both size and  beta porflio

\begin{center}
    \paperspacingnarrow
\input{Code/data/All_Industry_FF_pass_2}\\
\textbf{Table 7:} SIC top 10\% minus bottom 10\%
    \paperspacingwide
\end{center}

### 2021-05-02 Paper/data.tex
big push in latex

The data used for this paper was sourced from three datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily returns for the MSCI World ESG Focus Index and S\&P petroleum index, respectively. Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R. French - Data Library.

An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index. Petroleum companies act as a stand-in for a contra ESG portfolio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors. The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis. In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}. Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium. One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio. However \$Vice has only traded for two years, as such the data is far less robust. Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}. It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.

\begin{equation}
    \label{making_EMP}
    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
\end{equation}

For this analysis, daily returns data was collected between June 2011 and March 2021. The started was determined by the initiation of the MSCI ESG index. To collect daily individual stock return information CRSP data parsed from a larger data set and then collapsed on month by summing each individual days returns to create monthly returns. Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures. Finally the EMP factor was also  constructed summing the daily returns for each much. All three of the datasets were merged using the date and month as the matching component. The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security. Below in table 1 are the summary statistics for the raw data.

>>>>

The data used for this paper was sourced from four datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data for the daily returns of MSCI World ESG Focus Index and S\&P petroleum index was sourced from CAPIQ Compustat. Data on firm-level ESG scores (by ticker) was sourced from MSCI and reflected the most recent company ratings as MSCI does not have the historical ratings relatively available. Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R. French - Data Library.

The data and methodology for the paper are broken into two sections; first, a preliminary study is conducted using a broad base ESG factor to confirm a statistically risk sensitivity relationship between ESG and returns. Next, an analysis on a SIC level is conducted in an attempt to understand how industry-specific ESG factors affect the risk-return relationship of stock performance.

In the first stage of the analysis, an ESG factor is created by taking two market indexes and subtracting their daily returns in an attempt to isolate ESG based risk. The formula for this factor is represented in Equation \eqref{making_EMP}, and for the remainder of the section, one analysis will be referred to as EMP, which is sort for ESG minus petroleum. This method of creating a factor is very similar to Fama and French in their other four factors in the five-factor model. For this analysis, I will be using the MSCI World ESG Focus Index as the high ESG portfolio and the  S\&P petroleum index as the low ESG portfolio. The methodology for creating an ESG factor is similar to the work from \hl{put the author's name in here}. In \hl{put the author's name in here}'s work, they were able to find a statistical risk relationship between a high ESG portfolio and a petroleum index\cite{Frynas2005TheFD}. The petroleum index provides a good contra ESG portfolio as the petroleum industry is in direct opposition to environmental goals that are E components of ESG. A similar analysis could also be run with other market indexes. The S\&P Mining index is another popular index that also performs very poorly for environmental and social concerns\cite{Kapelus2002MiningCS}. A final index that also could represent a contra ESG portfolio is the  \$VICE ETF. This financial product tracks the market for  alcohol, drug, gaming, and betting industry which have all been shown as clear contra ESG industries \hl{inset citation}
\begin{equation}
    \label{making_EMP}
    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
\end{equation}

To conduct this analysis, daily data was collected between June 2011 and March 2021.  June 2011 marks the initiation of the MSCI ESG index; as such, this is the beginning of the sample period for this paper. Daily individual stock return information was collected from CRSP and merged with the daily ESG factor returns. The return data was collapsed on month to create a monthly return figure by summing all of the daily returns for each given month in the dataset. Next, the Fama French monthly factors were merged into the dataset for each individual firm for each month. After aggregating, the factor and individual firm return, there were 830,891 observations of monthly return data for 12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data for each individual security. Table 1 provides brief summary statistics for the data in the first stage of the analysis.

In the second stage of the analysis, I look at how SIC code-specific ESG factors affect the risk-return relationship. In this study, I used the MSCI ESG rating by ticker to created SIC-based ESG factors. MSCI rate companies on their ESG score AAA-CCC where AAA is the highest ESG rating and CCC is the lowest ESG rating. In the sample of MSCI ratings, 16 companies were AAA, 88 companies were AA, 118 companies were A, 170 companies were BBB, 108 companies were BB, 40 companies were B, and 5 companies earned a CCC rating.  Each of these ratings was assigned a mating numeric score (1-7) and were merged using the CRSP name file to their corresponding kypermno value by matching ticker information. Subsequently, the ESG scores were merged into the monthly return dataset described at the end of the first analysis. A final process was conducted to make a unique ESG factor for each of the SIC codes. The unique ESG factor was created at the first digit SIC level and generated by splitting the SIC level ESG data into deciles and taking the top 10\% and as the high ESG index bottom 10\% as the low ESG and subtracting high ESG minus low ESG as explained in equation \eqref{making_EMP_section}. This ESG factor was created for SIC codes beginning with 1-4 and 6-8. SIC code. SIC code 5 only had BBB ratings, and MSCI did not have significant data on SIC code 0 or 9. In addition to creating industry specific ESG factors I also created a all industry ESG factor using the top 10\% and bottom 10\% for all MSCI ESG scores to create an ESG factor. \hl{explain how EMP is the same down here for consistency of reuslts}

\begin{equation}
    \label{making_EMP_section}
    EMP_i = R_{\text{high ESG,i}}- R_{\text{low ESG,i}}
\end{equation}

\begin{center}
    Where i is the first digit SIC code    
\end{center}
